---
title: "Project1"
author: "Amirsaman"
date: "January 15, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# What is the total number of steps taken per day?

```{r echo=TRUE, results='asis'}
data <- read.csv("C:/Users/Bitcenter1/Dropbox/Coursera/Reproducable Research/Project1/activity.csv")
uniquedates <- unique(data$date)

TotalSteps <- rep(1,times=length(uniquedates))
MeanSteps <- rep(1,times=length(uniquedates))
MedianSteps <- rep(1,times=length(uniquedates))

for (i in 1:length(uniquedates)){
  steps <- data[data$date==uniquedates[i],1]
  TotalSteps[i] <- sum(steps,na.rm = TRUE)
  remove(steps)
}
hist(TotalSteps, main="Histogram of Total Steps per day",xlab = "Total Steps")
MeanSteps <- mean(TotalSteps,na.rm = TRUE)
MedianSteps <- median(TotalSteps,na.rm = TRUE)
```
**3. The mean of the total number of steps taken per day is `r MeanSteps` and the median is `r MedianSteps`.**

# what is the average daily activity pattern?
```{r echo=TRUE, results='asis'}
MeanStepsInterval <- rep(1,times=length(unique(data$interval)))
for (i in 1:length(unique(data$interval))){
  MeanStepsInterval[i] <- mean(data[data$interval==unique(data$interval)[i],1],na.rm = TRUE)
}
plot(unique(data$interval),MeanStepsInterval,type = "l",main="Time Series of average number of steps across days in time intervals",xlab = "Time Interval",ylab = "Average Number of Steps Across days")
MaxInt <- which.max(MeanStepsInterval)
```
**2.Maximum number of steps happens at `r MaxInt`th interval.**

#Imputing missing values
```{r echo=TRUE}
NumberOfNA <- sum(is.na(data[,1]))
```
**1. Total number of NAs in the data set is `r NumberOfNA`.**

**2 & 3. Filling missing values with average of the interval in a new data set called "Newdata".**
```{r echo=TRUE, results='asis'}
Newdata <- read.csv("C:/Users/Bitcenter1/Dropbox/Coursera/Reproducable Research/Project1/activity.csv")
uniquedates <- unique(Newdata$date)
NewMeanStepsInterval <- rep(1,times=length(unique(Newdata$interval)))
for (i in 1:length(unique(Newdata$interval))){
  NewMeanStepsInterval[i] <- mean(Newdata[Newdata$interval==unique(Newdata$interval)[i],1],na.rm = TRUE)
}
for (i in 1:length(Newdata[,1])){
  if (is.na(Newdata[i,1])){
    Newdata[i,1] <- NewMeanStepsInterval[match(Newdata[i,3],unique(Newdata$interval))]
  }
}
TotalStepsNew <- rep(1,times=length(uniquedates))
MeanStepsNew <- rep(1,times=length(uniquedates))
MedianStepsNew <- rep(1,times=length(uniquedates))

for (i in 1:length(uniquedates)){
  stepsNew <- Newdata[Newdata$date==uniquedates[i],1]
  TotalStepsNew[i] <- sum(stepsNew,na.rm = TRUE)
  remove(stepsNew)
}
hist(TotalStepsNew, main="Histogram of Total Steps per day for the new dataset")
MeanStepsNew <- mean(TotalStepsNew,na.rm = TRUE)
MedianStepsNew <- median(TotalStepsNew,na.rm = TRUE)
```
**4. The mean of the total number of steps taken per day is `r MeanStepsNew` and the median is `r MedianStepsNew`. These values are different from the first part. The imputation makes the data more symmetry.**

#Are there differences in activity patterns between weekdays and weekends?
```{r echo=TRUE}
Newdata <- read.csv("C:/Users/Bitcenter1/Dropbox/Coursera/Reproducable Research/Project1/activity.csv")
uniquedates <- unique(Newdata$date)
NewMeanStepsInterval <- rep(1,times=length(unique(Newdata$interval)))
for (i in 1:length(unique(Newdata$interval))){
  NewMeanStepsInterval[i] <- mean(Newdata[Newdata$interval==unique(Newdata$interval)[i],1],na.rm = TRUE)
}
for (i in 1:length(Newdata[,1])){
  if (is.na(Newdata[i,1])){
    Newdata[i,1] <- NewMeanStepsInterval[match(Newdata[i,3],unique(Newdata$interval))]
  }
}
Newdata$date <- as.Date(Newdata$date)
for (i in 1:length(Newdata$date)){
  if (is.element(weekdays(Newdata$date[i]), c("Saturday","Sunday"))){
    Newdata$Weekends[i] <- "Weekend"
  }
  else{
    Newdata$Weekends[i] <- "Weekday"
  }
}
Newdata$Weekends <- as.factor(Newdata$Weekends)
NewdataWE <- subset(Newdata,Newdata$Weekends=="Weekend")
NewdataWD <- subset(Newdata,Newdata$Weekends=="Weekday")


MeanStepsIntervalWE <- rep(1,times=length(unique(NewdataWE$interval)))
for (i in 1:length(unique(NewdataWE$interval))){
  MeanStepsIntervalWE[i] <- mean(NewdataWE[NewdataWE$interval==unique(NewdataWE$interval)[i],1],na.rm = TRUE)
}

MeanStepsIntervalWD <- rep(1,times=length(unique(NewdataWD$interval)))
for (i in 1:length(unique(NewdataWD$interval))){
  MeanStepsIntervalWD[i] <- mean(NewdataWD[NewdataWD$interval==unique(NewdataWD$interval)[i],1],na.rm = TRUE)
}
layout(matrix(1:2,ncol=1),widths=1,heights=c(2,2),respect=FALSE)
par(mar = c(0.5, 0.5, 1.5, 1.5), oma=c(5, 5, 1, 1), las=2)
plot(unique(NewdataWE$interval),MeanStepsIntervalWE,type = "l",main = "Weekends", xaxt='n',col="blue")
plot(unique(NewdataWD$interval),MeanStepsIntervalWD,type = "l",main = "Weekdays",col="blue")

mtext("x-axis", 1, 3, outer=TRUE)
mtext("y-axis", 2, 3, outer=TRUE, las=0)
```



